Nail: A practical interface generator for data formats

### 2014-04-01 paper/intro.tex
intro and title

Third, data formats can have structural dependencies, such as offset, length and checksum fields
that are hard to express in state-of-the-art grammar languages. Often the programmer is required to
manually write control code, such as re-positioning the parser's input stream, looping over the
invocation of a sub-parser or computing a
checksum over raw input bytes.  Besides leaving much room for errors with offset arithmetic, such
code is usually not reusable when generating output.

This paper presents the initial design and implementation of Nail, a
parser generator that greatly reduces the programmer effort required to
use a grammar-based parser.  Nail addresses the above three challenges
with several key ideas.  First, Nail XXX todo.  Bijections.  Reduced
expressiveness.  Constraints.  External interaction (checksum, ports,
etc).

We have implemented an initial prototype of Nail in C\@.  Our experience
so far suggests that it is a promising approach: we were able to construct
a succinct grammar for DNS packets, and write a small DNS server that
uses Nail for all packet input and output, and operates purely on
Nail-generated data structures, with no need for any semantic actions.

>>>>

Third, data formats, especially binary formats such as PNG or PDF, can have \emph{structural dependencies}, such as offset, length and checksum fields
that are hard to express in state-of-the-art grammar languages. Often the programmer is required to
manually write control code, such as re-positioning the parser's input stream, looping over the
invocation of a sub-parser or computing a
checksum over raw input bytes.  Besides leaving much room for errors with offset arithmetic, such
code is usually not reusable when generating output.

This paper presents the initial design and implementation of Nail, a parser
generator that greatly reduces the programmer effort required to use a
grammar-based parser. Nail addresses the above three challenges with several key
ideas. First, Nail reduces the expressiveness of its grammar language by
removing semantic actions. Existing parser generators allow arbitrary
computation to transform between the AST and the parser output.%  Instead, Nail
% derives the structure of its output automatically from the grammar, forcing the
% programmer to clearly separate syntactic validation and semantic processing.

Second, this allows the parser to construct a \emph{syntactic bijection} between input
and output. That is, parsing is the inverse of generating output. However,
Nail allows \emph{constants} in the external format to have multiple
representations if they should not affect the semantics of the data. For
example, in a text protocol, the amount of white-space separating tokens should
not affect the meaning of the data and is consequently not exposed to the
application by Nail in the internal representation. However, as long as
constants are only used for their intended purpose of representing syntax-only
features, the generated output will have the same semantics as the parsed input.

Third, Nail provides support for structural dependencies, transparently handling
offset and length fields. XXX todo

%  XXX todo.  Bijections.  Reduced
% expressiveness.  Constraints.  External interaction (checksum, ports,
% etc).

We have implemented an initial prototype of Nail for C\@.  Our experience
so far suggests that it is a promising approach: we were able to construct
a succinct grammar for DNS packets, and write a small DNS server that
uses Nail for all packet input and output, and operates purely on
Nail-generated data structures, with no need for any semantic actions.

### 2014-04-01 paper/main.tex
some typesetting tweaks
affiliation
acks
fixing up some bib stuff
kill another unused .bib file
stub conclusion
split off relwk
intro and title

\documentclass[10pt]{article}
\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{underscore}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{fullpage}

\begin{document}
\author{Julian Bangert and Nickolai Zeldovich}
\date{}
\title{Nail: A practical interface generator for binary formats}

@@

\input{abstract}
\input{background}
\input{design}
\input{impl}
\input{eval}
\input{future}

\bibliography{main,locasto,login,phrack,short,langsec}
\bibliographystyle{plain}

>>>>

\documentclass[10pt,twocolumn]{article}
\usepackage[10pt,nocopyright]{sigmin}
\usepackage[letterpaper, breaklinks, pdfborder={0 0 0}]{hyperref}
\usepackage[square,comma,numbers,sort&compress]{natbib}
\usepackage{times}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{underscore}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{fullpage}

\renewcommand{\ttdefault}{pxtt}   %% better \tt font

%% Ensure ligatures (e.g., ``fine official flag'') can be copy/pasted from PDF.
\input{glyphtounicode}
\pdfgentounicode=1

\begin{document}
\author{Julian Bangert and Nickolai Zeldovich \\ \em MIT CSAIL}
\date{}
\title{Nail: A practical interface generator for data formats}

@@

\input{abstract}
\input{intro}
\input{relwk}
\input{design}
\input{impl}
\input{eval}
\input{future}
\input{concl}
\input{ack}

\bibliography{n-str,main,locasto,phrack,langsec,n,n-conf}
\bibliographystyle{plain}
